Everthing you wanted to know about PanDA but were truly truly afraid to ask

### 2016-09-13 paper.tex
Add usual comments and headers
Itemize -> Enumerate
adding olcf boilerplate to acks
Second draft, plus integration of Sergey's part

@@

@@

\begin{abstract}
The abstract goes here. DO NOT USE SPECIAL CHARACTERS, SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT.
\end{abstract}

@@

% ----------------------------------------------------------------------------
% I
% ----------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

\begin{itemize}
  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
  \item Why are WMS needed?
  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
  \item \ldots
  \item Overview/Summary of paper.
\end{itemize}

@@

\begin{itemize}
  \item Brief overview of distributed CI.
  \item Workload vs Workflow
  \item Distinguish between WFMS and WLMS!!!!
  \begin{itemize}
    \item Role of Workload MS: help “hide” and federate heterogeneous resources.
  \end{itemize}
\end{itemize}

% ----------------------------------------------------------------------------
% III
% ----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}

\begin{itemize}
  \item Other workload management systems (WMS) (not exclusive)
  \begin{itemize}
    \item glidein-WMS
    \item Dirac
    \item ALICE-AlieN
  \end{itemize}
  \item Comparison and contrast
  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
\end{itemize}

% ----------------------------------------------------------------------------
% IV
% ----------------------------------------------------------------------------
\section{PanDA Overview}
\label{sec:panda_overview}

\begin{itemize}
  \item PanDA WMS
  \begin{itemize}
    \item State Diagrams
    \item Control Flow Diagrams
    \item Software Module and Architecture diagrams.
  \end{itemize}
  \item PANDA Pilot
\end{itemize}

@@

\begin{itemize}
  \item Challenges
  \item Opportunities
  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
  \item Take-aways, lessons learned
\end{itemize}

@@

\begin{itemize}
  \item Internal PanDA performance
  \item External PanDA performance (workload specific on underlying system/platform)
\end{itemize}

@@

\begin{itemize}
  \item Metadata performance issue, and how it’s resolved
  \item File I/O performance and impact
  \item ?
\end{itemize}

@@

% section panda_future_roadmap (end)

% \subsubsection{Subsubsection Heading Here}
% Subsubsection text here.

% \section{Type style and Fonts}
% Wherever Times is specified, Times Roman or Times New Roman may be used. If
% neither is available on your system, please use the font closest in
% appearance to Times. Avoid using bit-mapped fonts if possible. True-Type 1
% or Open Type fonts are preferred. Please embed symbol fonts, as well, for
% math, etc.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

% \section{Conclusion}
% The conclusion goes here. this is more of the conclusion

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank... More thanks here...

>>>>

\input{head}
\input{include}

@@

\newcommand{\sarp}[1]{{\color{green}\textit{sarp: {#1}}}}

@@

\begin{abstract}
 Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.  Heterogeneous  re- sources are distributed worldwide, thousands of physicists analyz- ing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year. The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment. In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model. The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.  PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
2013. We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
\end{abstract}

@@

% ----------------------------------------------------------------------------
% I
% ----------------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
The Large Hadron Collider (LHC) is created  to explore the fundamental  properties  of matter for the next decades. Since LHC start-up in 2009, the experiments has collected and distributed hundreds  of  petabytes  of  data worldwide to hundreds of computer centers. Thousands of physicists analyze peta-scale  data volumes daily. One of  the LHC experiments  the ATLAS  [1]  utilizes the PanDA  workload management system [4] (WMS) for distributed data processing and analysis. The ATLAS Computing model [2] is based on a Grid paradigm [3], with multilevel, hierarchically distributed computing  and storage resources. Acronym  PanDa stands for Production  and Distributed Analysis, the system  has  been developed to meet growing ATLAS production and analysis requirements for a data-driven  workload  management system capable of operating at LHC data processing scale.
PanDA has a highly scalable architecture. Scalability  has been demonstrated in ATLAS through the rapid increase in usage over the past several years of operations, and is expected to meet the continuously growing number of jobs over the next decade. PanDA  was designed to have the flexibility to adapt to emerging computing  technologies in processing,  storage, networking  as well as the underlying software stack (middle- ware). This flexibility has also been successfully demonstrated through the past six years of evolving  technologies adapted by computing centers in ATLAS which span many continents and yet are seamlessly integrated into PanDA.
PanDA  manages a wide spectrum of workloads, ranging from raw data processing to Monte Carlo simulation and user anal- ysis, while actively evolving to meet rapidly changing science needs.  Today, it  serves several  thousand users, successfully managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU cores  which process  more than a million jobs per day, Figure 1 shows  a  summary   of daily completed jobs on ATLAS Grid for the past 12 month.
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figure/DailyJobs.png}
\caption{Daily completed jobs on ATLAS Grid\label{fig:daily}}
\end{center}
\end{figure}

-------------------------------------------------------------------------------------------
\begin{enumerate}
  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
  \item Why are WMS needed?
  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
  \item \ldots
  \item Overview/Summary of paper.
\end{enumerate}

@@

\begin{enumerate}
  \item Brief overview of distributed CI.
  \item Workload vs Workflow
  \item Distinguish between WFMS and WLMS!!!!
  \begin{enumerate}
    \item Role of Workload MS: help “hide” and federate heterogeneous resources.
  \end{enumerate}
\end{enumerate}

% ----------------------------------------------------------------------------
% III
% ----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}
\input{WMSLiterature}
--------------------------------------------------------------------------------------------------
\begin{enumerate}
  \item Other workload management systems (WMS) (not exclusive)
  \begin{enumerate}
    \item glidein-WMS
    \item Dirac
    \item ALICE-AlieN
  \end{enumerate}
  \item Comparison and contrast
  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
\end{enumerate}

% ----------------------------------------------------------------------------
% IV
% ----------------------------------------------------------------------------
\section{PanDA Overview}\label{sec:panda_overview}
PanDA consists of several interconnected subsystems, most of them are built from off-the-shelf and Open Source components. Figure \ref{fig:architecture} shows a schematic view of the PanDA system.
In the following we will briefly describe system’s architecture and components.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figure/PandaArch.jpg}
\caption{Schematic view of the PanDA system\label{fig:architecture}}
\end{center}
\end{figure}
\subsection{PanDA Server}
The PanDA server is the main component of the system, It provides  a task queue managing all job information centrally. The PanDA  server receives jobs through the client interface into the task queue, upon which a brokerage module  operates to prioritize  and assign work on the basis of job type, priority, input data and its locality, and available CPU resources. The PanDA  server operates as a web service. It runs on Apache web server, interacting with the back-end database running  on separate servers. It uses the Apache worker  model, with many independent   processes  handling client requests  in  parallel. Since all  state is maintained  in  the central database,  the PanDA Server application instance itself is stateless. Currently production version of PanDA utilizes Oracle database backend, but PanDA  can also work with MySQL family of databases.

\subsection{PanDA Pilot}
 PanDA is a pilot based workload  management system. In the PanDA job lifecycle, pilot jobs (Python scripts that organize workload processing on a worker  node) are submitted to sites. When these  pilot jobs start on a  worker node they contact a central  server  to retrieve  a real payload (i.e., an end-user job) and execute it. Using these pilot-based  workflows helps to improve job reliability, optimize resource utilization, allows for opportunistic  resources usage, and mitigates  many of the problems  associated with the inhomogeneities found on the Grid.
The pilots themselves  do not contain all the functionality needed to request a payload  job from the Panda Server. First, the complete  set of pilot code is downloaded via HTTP from a  central Subversion repository.  The repository works with Apache web server configured  with a  memory-based   web proxy (Squid). The purpose  of the cache  is to reduce  the

request load on the back-end Subversion  server. That allows for a  very good performance since only occasional  queries trigger a  full  lookup on the back-end  Subversion  system, and most external  queries are pulled from memory on the front-end web server. The benefits of this system include a high-performance code download service combined with code updates still being immediately  available  as soon as they are committed to source code control.

\subsection{PanDA Factory}
As a pilot-based  system, PanDA  requires some way to get the initial PanDA pilot onto worker nodes at sites. This is done with the help of a component  called AutoPyFactory (APF). APF runs in a single daemonized process, launching  a separate thread for each internal  workflow. Each one of these internal workflows typically serves  a single job queue  as defined  in WMS, and delivers pilots to a single batch queue, either local or remote. The behavior of these APF workflows is determined by the combination of a set  of plug-ins, invoked in a  fixed order, in a loop, each one in charge of the performance of a well defined action.

\subsection{PanDA Monitoring}
The PanDA Monitor is a web-based dashboard-style  graphi- cal application. It runs on Apache and interacts with the back- end database for persistence. The Monitor is the primary way that users and site administrators get a  view into the status of current and past jobs, data movement,  pilot factory. The Monitor allows users  access  to all job log files on job by job basis thus greatly simplifying code debugging and failure analysis in distributed computing environment.

\section{Evolution of PanDA WMS}
While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide. The Worldwide LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned analysis and data processing, but it will be insufficient  for Monte Carlo (MC) production and any extra activities. Additional computing and storage  resources are therefore required.  To alleviate  these challenges, ATLAS is engaged  in an ambitious  program to expand  the current computing model to include  additional  resources such as the opportunistic   use of supercomputers   as well as commercial and academic clouds. In turn this activity drives evolution of the PanDA WMS.

\subsection{Use of Supercomputers with PanDA}
The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers. Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network. The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors. On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI). While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization. It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run. What matters is the total number of cycles that can be offloaded from the Grid. Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc. A reorganization  of the standard  workflow is therefore needed.
The most suitable task for HPCs in HEP is event generation. Event generators are mostly computational,  stand-alone code, with few input requirements.  There is no need  to stage-in much data. The event generation  in ATLAS corresponds  to
10-15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs is detector  simulation. Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.  However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
\subsection{Interfacing PanDA with Titan}
The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US. Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
Integration with Titan is the current focus for PanDA develop- ers. The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface [6]. This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world. The pilot can communicate with the PanDA server from the front-end machine. The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job. ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design. Time allocation on an HPC is competitive and large projects are often preferred. About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources. The worker nodes
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figure/PandaInterfaceWithHPC.png}
\caption{Schematic view of the PanDA system\label{fig:architecture}}
\end{center}
\end{figure}

-------------------------------------------------------------------------------------
\begin{enumerate}
  \item PanDA WMS
  \begin{enumerate}
    \item State Diagrams
    \item Control Flow Diagrams
    \item Software Module and Architecture diagrams.
  \end{enumerate}
  \item PANDA Pilot
\end{enumerate}

@@

\begin{enumerate}
  \item Challenges
  \item Opportunities
  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
  \item Take-aways, lessons learned
\end{enumerate}

@@

\begin{enumerate}
  \item Internal PanDA performance
  \item External PanDA performance (workload specific on underlying system/platform)
\end{enumerate}

@@

\begin{enumerate}
  \item Metadata performance issue, and how it’s resolved
  \item File I/O performance and impact
  \item ?
\end{enumerate}

@@

 \section{Conclusion}\label{sec:conclusion}
The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.  In the process,  the old batch job paradigm  of
computing in HEP was  discarded  in favor  of a  far more flexible and scalable  model. The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments. PanDA  is the first exascale  workload manage- ment system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013. Next LHC run will pose massive computing  challenges. With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more. Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS. To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system. Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry. The technology base provided by the PanDA system will enhance the usage of a variety  of high- performance computing resources available to basic research.

\input{ack}

%\section*{Acknowledgment}

%The authors would like to thank... More thanks here...
